---
title: "Exploring-temperature-anomalies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{exploring-temperature-anomalies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

This vignette shows how to explore `temp_data` (annual temperature anomaly vs the 1861–1890 baseline) and reproduce the main figure used in the app.

## Data

`temp_data` has three columns:

- `country` — region name (e.g., "World", "Northern Hemisphere")
- `year` — observation year
- `temp_anomaly` — °C relative to the 1861–1890 baseline

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(bushfireApp)
library(ggplot2)
```


```{r}
head(temp_data)
```

```{r}
#| fig.cap: "Temperature anomaly over time"
#| fig.alt: "A time-series line chart of temperature anomaly relative to 1861–1890 baseline."

ggplot(temp_data, aes(x = year, y = temp_anomaly, colour = country)) +
geom_line(linewidth = 1) +
geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
labs(
x = "Year",
y = "Temperature anomaly (°C vs 1861–1890)",
title = "Temperature anomaly over time"
) +
theme_minimal(base_size = 12)
```

## How to interpret
- Values above 0 are warmer than the 1861–1890 baseline; below 0 are cooler.

- A persistent upward trend indicates warming.

- Recent decades show sustained positive anomalies, consistent with modern warming.

## Reproducibility

Source: Our World in Data “temperature-anomaly” (subset cached in the package). Data pipeline in data-raw/build_data.R.
